Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-19 paper/agree.tex
revised content added to semantics section. Checkpointing

This section formalizes AGREE's assume-guarantee reasoning by defining the set of verification conditions that AGREE must check.
The verification conditions succinctly summarize what in means when AGREE proves that a system implementation satisfies it specification.
These conditions are useful in understanding how the semantics in synthesis differ from those in the assume-guarantee reasoning in \secref{sec:semantics}.

This section then introduces the AGREE contract specification language through the cyber-hardened system of \figref{fig:hardened}.
It details the specification of that top-level system interface.
It then discusses the specifications for the filter and the monitor.
These two specifications are those that are synthesized to equivalent CakeML by SPLAT.

>>>>

This section gives an overview of AGREE's assume-guarantee reasoning
system, abstractly defining the verification conditions that AGREE
checks in order to establish overall system correctness at the model
level.  The discussion then becomes more concrete, showing how the
AGREE contract specification language is used to specify the
cyber-hardened system of \figref{fig:hardened}. It explains the the
top-level specification of the example system, then shifts to the
specifications for the filter and the monitor.  These two
specifications are those that are synthesized to equivalent CakeML by
SPLAT.

%% These conditions are useful in understanding how the semantics in
%% synthesis differ from those in the assume-guarantee reasoning
%% in \secref{sec:semantics}.

### 2022-03-19 paper/high-assurance-contracts.tex
revised content added to semantics section. Checkpointing

As noted previously, the system with the untrusted AI does not pass
AGREE verification.  Transformations in BriefCASE cyber-harden the
system against malicious AI behavior by inserting a high-assurance
filter and monitor between the AI and the WM as shown in
\figref{fig:hardened}.  (In our context \emph{high-assurance} means
that the behaviors of the components are verified to meet their
contracts.)  In the example, the inserted filter protects the WM from
malformed data and the monitor protects the WM from (possibly malicious)
spontaneous or delayed responses.  These high-assurance components are
added, one at a time, by
\begin{itemize}
  \item selecting the connection in the model where the
    component is to appear,
  \item choosing the appropriate transformation, and
  \item specifying the relevant filtering or monitoring \emph{policy} (behavior).
\end{itemize}

>>>>

As noted previously, the system with the untrusted AI does not pass
AGREE verification.  Transformations carried out in the BriefCASE
interface cyber-harden the system against malicious AI behavior by
inserting a high-assurance filter and monitor between the AI and the
WM as shown in
\figref{fig:hardened}.  (In our context \emph{high-assurance} means
that the behaviors of the components are verified to meet their
contracts.)  In the example, the inserted filter protects the WM from
malformed data and the monitor protects the WM from (possibly malicious)
spontaneous or delayed responses.  These high-assurance components are
added, one at a time, by
\begin{itemize}
  \item selecting the connection in the model where the
    component is to appear,
  \item choosing the appropriate transformation, and
  \item specifying the relevant filtering or monitoring \emph{policy} (behavior).
\end{itemize}

### 2022-03-19 paper/semantics.tex
revised content added to semantics section. Checkpointing

Brief, high-level overview, of the AGREE semantics for component synthesis.
These semantics are those defined in HOL.
The component is defined by its state.
The state is its internal state and current output.
Input streams are constrained to obey the assumptions.
At each input, the \textbf{eq}-statements are evaluated, in order, to update the internal state and outputs.
After the \textbf{eq}-statements are evaluated, guarantees are checked.
The section should define clearly what it means for a component to be \emph{well-formed}. 
The section should conclude by stating the \emph{Component Correct} definition.

The correctness, and structure, of the synthesized CakeML code is based on these semantics.

\subsection{Old AGREE Semantics}
\label{agree-semantics}

Here the formal semantics of the AGREE contract specification are
briefly presented to make clear the meaning of a high-assurance
component.  These semantics are the basis of proofs showing that a
synthesized component has input to output behavior specified by its
contract.

Assume that all data flowing between components is in the form of byte
arrays, thus input and output is modelled at the level of what is
passed ``over the wire'' by the communication fabric. An \emph{environment},
$\theta: \lval \mapsto \konst{string}$, binds \emph{L-values} to
strings where an L-value is anything that can appear on the left-hand
side of an assignment such as a named port, a field in a record, a
entry in an array, a local value defined by an eq-statement etc.

Let $s$ be an AGREE contract specification for some high-assurance
component. The notation $\theta_s$ is used to denote the environment that
contains a binding for any L-value in the scope of $s$ and nothing else, and
the notation $\Theta_s$ denotes the universe of all such environments. The
semantics of $s$ are defined over streams, $\pi = \theta_1, \theta_2, \ldots$, which
are finite sequences of environments, $\pi \in \Theta_s^*$.

%% \begin{comment}
%% [\emph{So basically $\theta$ covers the inputs and the state, i.e. we can
%% generate a contig type for the inputs plus state vars.} ]

%% [Question: what is a \emph{total} environment? One that supplies at
%% least the bindings in $\theta$ needed to evaluate the spec?]
%% \end{comment}

The function $\konst{eval}\; s\; \pi$ evaluates $s$ on the stream,
$\pi$, and returns \konst{true} if $s$ is invariant along the entire
stream and \konst{false} otherwise. A guarantee $\mathcal{G}$ in $s$
is \emph{invariant} if $\mathcal{G}$ is true for each prefix of $\pi$,
while an eq-statement in $s$ is invariant if its binding in the
context of every step is equivalent to the computed value of its
associated expression in that same context. All guarantees and
eq-statements must be invariant in the stream for the function to
return \konst{true}.

The meaning of an AGREE contract specification is now defined as the
set of environment streams on which it is invariant.
%
\[
  \Lang{s} = \set{\pi \in \Theta_s^* \mid \konst{eval}\ s\ \pi = \konst{true}}
\]
%
\noindent Intuitively, any stream $\pi \in \mathcal{L}(s)$, at each step, binds
the L-values in the eq-statements in a way that is consistent with
their associated expressions and the guarantees are all true in that
same step.

We claim that a synthesized high-assurance component preserves the
input to output relationship in the specification $s$ over every
stream in $\Lang{s}$.  Let $\pi^\prime = \konst{SynthEval}\;
s\; \pi$ denote a function that synthesizes $s$ and then evaluates that
synthesized component on the stream $\pi$ to create a new stream
$\pi^\prime$ containing added output and other bindings. We say that
two streams are equivalent in regards to a specification, denoted as
$\pi =_s \pi^\prime$, if and only if the two streams are the same
length and agree on bindings for the input and output for $s$ at every
step of the streams.  We now formally state the correctness claim for
synthesis.
%
\[
  \forall \pi \in \Lang{s}.\; (\konst{SynthEval}\; s\; I_s(\pi)) =_s \pi\\
\]
%
where $I_s(\pi)$ returns the corresponding stream that only retains
bindings for inputs in each step and nothing else. The claim is that
the synthesized component generates the same output stream as the
specification for any stream belonging to the specification that is
restricted to just input bindings at each step.

>>>>

In this section we give an overview of the AGREE semantics, a
machine-checked formal definition of the meaning of AGREE expressions,
formulas, and contracts. The semantics supports a formal connection
between the results of the AGREE model checker and code generated from
specifications of leaf level cyber-components. It also provides the root
meaning at the base of a chain of translation steps moving from AGREE
specifications to CakeML executables.

\subsection{AGREE syntax and semantics}
\label{agree-semantics}

The syntax of AGREE is essentially that of quantifier-free first order
logic (FOL) supplemented with a few temporal operators. The terms
(\emph{e}) are the usual arithmetic expressions built from variables
and numeric literals, while formulas (\emph{b}) are built using the
standard logical connectives from atomic formulas (\emph{a}) based on
the usual comparison operators.
\[
\begin{array}{rcl}
e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
a & ::= & e\; \set{=,<,\leq}\; e \\
b & ::= & v \mid c \mid a \mid \neg b
            \mid b \; \set{\land,\lor,\imp,\iff}\; b
\end{array}
\]

There is a conditional, $\itelse{b}{(-)}{(-)}$, for both terms and
formulas. There are also temporal operators $\konst{pre}(-)$, delay
$(-) \to (-)$, and $\konst{Historically}(-)$.

The semantics of terms and formulas is in terms of \emph{streams of
values}. Values encompass at least booleans and numbers, but can be
readily extended to include records and arrays. A value stream is a
total function from natural numbers to values:
\[
 \konst{stream} = \mathbb{N} \to \konst{value}
\]
Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$
binding variable names to value streams, the semantics $\sem{-}$ of terms and
formulas, omitting temporal operators, is standard.

\[
\begin{array}{rcl}
\sem{v}_t & = & E(v)(t) \\
\sem{c}_t & = & c \\
\sem{e_1 + e_2}_t & = & \sem{e_1}_t + \sem{e_2}_t \\
   & \cdots & \\
\sem{b_1 \land b_2}_t & = & \sem{b_1}_t \land \sem{b_2}_t \\
   & \cdots & \\
\end{array}
\]
